---
title: "Lecture 7: Gelman Hill Ch 4.1 - 4.3"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(show.signif.stars = FALSE)
library(tidyverse) 
library(arm)
```


For various reasons, data transformations may be necessary or result in better interpretations for regression models.

### Linear Transformations

Linear transformations of predictors can be formulated as:

\vfill

Linear transformations of the predictors do not influence the fit or predictions of a regression model. 
\vfill

Recall the general interpretation of the regression coefficients is "the average difference in y when comparing units that differ by one unit, on predictor j, and are otherwise identical." However, consider two covariates:



\vfill

```{r, message = F}
library(readr)
library(arm)
Seattle <- read_csv('http://math.montana.edu/ahoegh/teaching/stat408/datasets/SeattleHousing.csv')
```

```{r}
summary(lm(price ~ bedrooms + sqft_living, data = Seattle))
```


\newpage
Furthermore, the interpretation of the intercept is still a little confusing. In this case, we are looking at a house with zero bedrooms and zero square feet of living space.
\vfill

#### Standardization

One common option is to standardize the predictors using a z-scale.



\vfill

```{r}
library(dplyr)
Seattle <- Seattle %>% mutate(sqft_z = (sqft_living - mean(sqft_living))/sd(sqft_living),
                              bedrooms_z = (bedrooms - mean(bedrooms))/sd(bedrooms))
```

Note it is important to interpret the zero values for each covariate, so the mean living space is `r round(mean(Seattle$sqft_living))` and the mean number of bedrooms is `r round(mean(Seattle$bedrooms),2)`.

\vfill

```{r}
lm_standard <- lm(price ~ bedrooms_z + sqft_z, data = Seattle)
summary(lm_standard)
```

\newpage

Now the intepretation of the parameters is:

- (Intercept): The predicted price of a house with 
\vfill

- (bedrooms_z): The average difference in price when comparing houses that differ by 
\vfill

- (sqft_z): The average difference in price when comparing houses that differ by 

\vfill

Note when summarizing the coefficients for homework, exams, or projects, make sure to talk about the size of the difference and include confidence intervals in the discussion.

```{r}
confint(lm_standard)
```

\vfill

The data can also be centered and/or standardized using different approaches. 

\vfill
\vfill
\vfill

\newpage

#### Correlation

Consider a regression line $y = \beta_0 + \beta_1 x$, where both $x$ and $y$ are standardized.
\vfill
```{r}
Seattle <- Seattle %>% mutate(y = (price - mean(price))/sd(price),
                              x = (sqft_living - mean(sqft_living))/sd(sqft_living))
library(ggplot2)
Seattle %>% ggplot(aes(y = y, x=x)) + geom_point(alpha = .1) + geom_smooth(method = 'lm') +
  ggtitle('Correlation between Price and Square Footage') + ylab('Price (standardized)') + xlab('Square Footage (standardized)') + annotate("text", x = 6, y = 2, label = paste("Slope is", round(cor(Seattle$y,Seattle$x),2))) + geom_segment(x = 6, y = 2.2, xend = 5, yend = 3.8, arrow = arrow(length = unit(.5,"cm")))

display(lm(y~x, data = Seattle))
cor(Seattle$y,Seattle$x)
```

\vfill
_Regression to the mean:_ 

\newpage
